---
title: "SEGUL performance benchmark"
author: Heru Handika
output: html_document
---

This file contains all code for Handika and Esselstyn (in review) manuscript. I use the R base pipe (required R version 4.1). If you are using R version <4.1, change the pipe "|>" to magrittr pipe "%>%".

```{r}
if (!require(pacman)) {
  install.packages(pacman)
}

pacman::p_load(here)
pacman::p_load(ggplot2)
pacman::p_load(readr)
pacman::p_load(dplyr)
pacman::p_load(stringr)
pacman::p_load(RColorBrewer)

source(here::here("R", "utils.R"))
```

## Load benchmark results for Handika & Esselstyn. In review.

The CSV files is generated using [benc-parser](https://github.com/hhandika/bench-parser). Below, we clean the file and change any values in zero to NA and write the cleaned data frame into a new CSV file.s

```{r}
file_path <- here("data", "data_ms_rev2.csv")
bench <- readr::read_csv(file_path) |>
  # ignore SEGUL GUI (Linux) Percent_CPU_usage. 
  # Not relevant for the paper and it does not seem to show the actual usage
  dplyr::mutate(Percent_CPU_usage = ifelse(Apps == "SEGUL GUI (Linux)", NA, Percent_CPU_usage)) |>
  dplyr::mutate(RAM_usage_Mb = na_if(RAM_usage_Mb, 0.0)) |>
  dplyr::mutate(Percent_CPU_usage = na_if(Percent_CPU_usage, 0.0)) |>
  # Latest bench is irrelevant since everything is the latest now.
  dplyr::select(-Latest_bench)

readr::write_csv(bench, here("results", "benchmark.csv"), na = "")
```

## Summarize benchmark results

We check the benchmark result. Latest benchmark only generate 5 replicates. Hence, we check if each datasets for each feature test contains 5 replicates. Any unmatched values may be a result of parsing errors or issues with the raw files generated in our automatic test bench. Go back to the raw files to find out about the errors.

The next step is to generate means and standard deviation for each dataset. Then, save the file as CSV.

```{r}
bench_results <- c("Execution_time_secs", "RAM_usage_Mb", "Percent_CPU_usage") 

# Get replicate counts
counts <- bench |>
  dplyr::count(Analyses, Datasets, Platform, OS_name, Apps)

# Check each contains 5 iterations
mean(counts$n)

summary_by_datasets <- bench |>
  dplyr::group_by(Analyses, Datasets, Platform, OS_name, Apps) |>
    dplyr::summarise(
    across(bench_results, ~ mean(.x, na.rm = TRUE), .names = "Mean_{.col}"),
    across(bench_results, ~ sd(.x, na.rm = TRUE), .names = "SD_{.col}"),
    )

readr::write_csv(summary_by_datasets, here("results", "mean_bench_by_dataset.csv"), na = "")
```

## Mean summary by analyses

We also generate summary per type of analyses (i.e., per benchmark test), per input format. We only need summary per analyses type as CSV that we use for the Table 4 in the manuscript. This comparison provide general performance of tested applications across features and input formats.

```{r}
analyses.df <- bench |>
  # Create a analyses type column by removing input format
  dplyr::mutate(AnalysesType = str_extract(Analyses, "^[^(]+")) 

# Calculate mean and sd for each column
summary_by_analyses_type <- analyses.df |>
  dplyr::group_by(AnalysesType, Apps) |>
  dplyr::summarise(
    across(bench_results, ~ mean(.x, na.rm = TRUE), .names = "Mean_{.col}"),
    across(bench_results, ~ sd(.x, na.rm = TRUE), .names = "SD_{.col}"),
    )
  
summary_by_analyses <- analyses.df |>
  dplyr::group_by(AnalysesType, Datasets, Apps) |>
  dplyr::summarise_at(bench_results, mean)


summary_by_input_format <- analyses.df |>
  # Create a input format column by removing analyses names
  dplyr::mutate(format = str_extract(Analyses, "(?<=\\()([^)]+)(?=\\))")) |>
  dplyr::group_by(AnalysesType, format, Platform, OS_name, Apps) |>
  dplyr::summarise_at(bench_results, mean)

# Only summary by type is needed. The rest is for exploratory only.
readr::write_csv(summary_by_analyses_type, here("results", "mean_bench_by_analyses.csv"), na = "")
```


## Generate plots using the mean of benchmark results for each datasets

This result is for Figure 1. We exclude the read summary. We used it separately for Figure 2. 

```{r}
summary_by_datasets |>
  dplyr::filter(OS_name == "Linux") |>
  dplyr::filter(Analyses != "Read Summary (FASTQ)") |>
  ggplot(aes(x = Mean_Execution_time_secs, y = Mean_RAM_usage_Mb)) +
    geom_point(size = 3, aes(color = Datasets, shape = Apps)) +
    facet_wrap(~ Analyses, scales = "free", ncol = 3) +
    scale_color_manual(values = ai_2024, guide = guide_legend(nrow = 6)) +
    scale_shape_manual(values = point_shapes, guide = guide_legend(nrow = 5)) +
    coord_trans(x = "log") +
    scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(x = "Mean execution time (secs)", y = "Mean RAM usage (Mb)")

.height <- 9
.width <- 10

ggsave(here("figures", "summary_fig1.png"), width = .width, height = .height, units = "in")
ggsave(here("figures", "summary_fig1.pdf"), width = .width, height = .height, units = "in")
```
## Figure 2 SEGUL GUI versus CLI performance for genome analyses

```{r}
bench |>
  dplyr::filter(OS_name == "Linux") |>
  dplyr::filter(Analyses == "Read Summary (FASTQ)") |>
  dplyr::mutate(Apps = ifelse(Apps == "SEGUL GUI (Linux)", "SEGUL GUI", Apps)) |>
  ggplot(aes(x = Execution_time_secs, y = RAM_usage_Mb)) +
    geom_point(size = 3, aes(color = Apps, shape = Apps)) +
    scale_color_manual(values = ai_2024) +
    scale_shape_manual(values = point_shapes) +
    coord_trans(x = "log") +
    scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
    theme_minimal() +
    labs(x = "Execution time (secs)", y = "Memory usage (Mb)")

.height <- 3
.width <- 4

ggsave(here("figures", "summary_fig2.png"), width = .width, height = .height, units = "in")
ggsave(here("figures", "summary_fig2.pdf"), width = .width, height = .height, units = "in")
```


## Figure 3 Execution time comparison across different SEGUL

```{r}
apps <- c("SEGUL CLI", "SEGUL GUI (Linux)", "SEGUL GUI (Android)", "SEGUL GUI (iPadOS)")
analyses <- c("Alignment Concatenation (NEXUS)", "Alignment Conversion (NEXUS)", "Alignment Splitting (NEXUS)", "Alignment Summary (NEXUS)", "Sequence Removal (NEXUS)")
dataset <- c("Oliveros et al. 2019 (522.5 MBases, DNA)", "Shen et al. 2018 (398.8 MBases, AA)")

segul_compare <- bench |>
  dplyr::filter(Apps %in% apps) |>
  dplyr::filter(Analyses %in% analyses) |>
  dplyr::filter(Datasets %in% dataset) |>
  # Both dataset will be NEXUS, so don't need to display it in the plot.
  dplyr::mutate(Analyses = str_remove_all(Analyses, "\\(NEXUS\\)")) |>
  dplyr::mutate(Apps = ifelse(Apps == "SEGUL CLI", "SEGUL CLI (Linux)", Apps))

segul_compare_by_analyses <- segul_compare |>
  dplyr::group_by(Analyses, Apps, Datasets) |>
  dplyr::summarise(
    mean_execution = mean(Execution_time_secs),
    sd_execution = sd(Execution_time_secs),
    lower = mean_execution - sd_execution,
    upper = mean_execution + sd_execution
  )

segul_compare_by_analyses |>
  ggplot(aes(x = Analyses, y = mean_execution, fill = Apps)) +
    geom_bar(stat = "identity", position=position_dodge()) +
    geom_errorbar(aes(ymin = lower, ymax = upper), position=position_dodge()) +
    facet_wrap(~ Datasets,  ncol = 4) +
    scale_fill_manual(values = ai_2024_4) +
    scale_x_discrete(guide = guide_axis(angle = 60)) +
    theme_minimal() +
    labs(y = "Mean execution time (secs)")


.height <- 6
.width <- 8

ggsave(here("figures", "summary_fig3.png"), width = .width, height = .height, units = "in")
ggsave(here("figures", "summary_fig3.pdf"), width = .width, height = .height, units = "in")
```
## Extra figures for documentation

```{r}
analyses <- c("Alignment Concatenation (NEXUS)")
apps <- c("SEGUL CLI", "AMAS", "SEGUL GUI (Linux)", "SEGUL CLI (--datatype ignore)", "SEGUL GUI (Android)", "SEGUL GUI (iPadOS)", "goalign (single-core)")
apps_mem_usage <- c("SEGUL CLI", "AMAS", "SEGUL GUI (Linux)", "SEGUL CLI (--datatype ignore)", "goalign (single-core)")
dataset <- c("Oliveros et al. 2019 (522.5 MBases, DNA)", "Shen et al. 2018 (398.8 MBases, AA)")

segul_compare_docs <- bench |>
  dplyr::filter(Apps %in% apps) |>
  dplyr::filter(Analyses == "Alignment Concatenation (NEXUS)") |>
  dplyr::filter(Datasets %in% dataset) |>
  # Both dataset will be NEXUS, so don't need to display it in the plot.
  dplyr::mutate(Analyses = str_remove_all(Analyses, "\\(NEXUS\\)"))

segul_compare_by_analyses_docs <- segul_compare_docs |>
  dplyr::group_by(Analyses, Apps, Datasets) |>
  dplyr::summarise(
    mean_execution = mean(Execution_time_secs),
    sd_execution = sd(Execution_time_secs),
    mean_ram_usage = mean(RAM_usage_Mb),
    sd_ram_usage = sd(RAM_usage_Mb),
    lower = mean_execution - sd_execution,
    upper = mean_execution + sd_execution,
    lower_ram_usage = mean_ram_usage - sd_ram_usage,
    upper_ram_usage = mean_ram_usage + sd_ram_usage
  )

# Execution time
segul_compare_by_analyses_docs |>
  dplyr::mutate(AppGroup = ifelse(str_detect(Apps, pattern="SEGUL"), "SEGUL", Apps)) |>
  ggplot(aes(x = Apps, y = mean_execution, fill = AppGroup)) +
    geom_bar(stat = "identity", position=position_dodge()) +
    geom_errorbar(aes(ymin = lower, ymax = upper), position=position_dodge()) +
    facet_wrap(~ Datasets,  ncol = 4) +
    scale_fill_manual(values = ai_2024_4) +
    scale_x_discrete(guide = guide_axis(angle = 60)) +
    theme_minimal() +
    theme(legend.position = "none") +
    labs(y = "Mean execution time (secs)")

ggsave(here("figures", "summary_docs1.png"), width = .width, height = .height, units = "in")
ggsave(here("figures", "summary_docs1.pdf"), width = .width, height = .height, units = "in")

# Execution time
segul_compare_by_analyses_docs |>
  dplyr::filter(Apps %in% apps_mem_usage) |>
  ggplot(aes(x = Apps, y = mean_ram_usage, fill = Datasets)) +
    geom_bar(stat = "identity", position=position_dodge()) +
    geom_errorbar(aes(ymin = lower_ram_usage, ymax = upper_ram_usage), position=position_dodge()) +
    facet_wrap(~ Datasets,  ncol = 4) +
    scale_fill_manual(values = ai_2024_4) +
    scale_x_discrete(guide = guide_axis(angle = 60)) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(y = "Mean RAM usage (Mb)")


.height <- 4
.width <- 6

ggsave(here("figures", "summary_docs2.png"), width = .width, height = .height, units = "in")
ggsave(here("figures", "summary_docs2.pdf"), width = .width, height = .height, units = "in")
```

