---
title: "SEGUL performance benchmark"
author: Heru Handika
output: html_document
---

This file contains all code to analyses SEGUL benchmark results. I use the R base pipe, required R version 4.1. If you are using R version <4.1, change the pipe "|>" to magrittr pipe "%>%".

```{r}
library(here)
library(ggplot2)
library(readr)
library(dplyr)
library(stringr)
library(RColorBrewer)
```


```{r}
bench <- readr::read_csv(here("data", "benchmark_results.csv")) |>
  dplyr::mutate_at("CPU_usage", stringr::str_remove_all, "%") |>
  dplyr::mutate_at("CPU_usage", readr::parse_number)

latest_bench  <- bench |> 
  dplyr::filter(Latest_bench == TRUE)
```

## Summarize benchmark results

```{r}
bench_results <- c("Execution_time_secs", "RAM_usage_Mb", "CPU_usage")

latest_bench.pub <- latest_bench |>
  dplyr::filter(Apps != "Phyluce") |>
  dplyr::filter(Analyses != "Alignment splitting")

# Check the benchmark for each dataset contains 10 iterations
counts <- latest_bench.pub |>
  dplyr::count(Analyses, Datasets, OS_name, Apps)

mean(counts$n)

# Generate mean of each benchmark result

summary_by_os <- latest_bench.pub |>
  dplyr::group_by(Analyses, OS_name, Apps) |>
  dplyr::summarise_at(bench_results, mean)


readr::write_csv(summary_by_os, here("data", "mean_bench_by_os.csv"))

summary_by_analyses <- latest_bench.pub |>
  dplyr::group_by(Analyses, Apps) |>
  dplyr::summarise_at(bench_results, mean)


readr::write_csv(summary_by_analyses, here("data", "mean_bench_by_analyses.csv"))

summary_by_datasets <- latest_bench.pub |>
  dplyr::group_by(Analyses, Datasets, OS_name, Apps) |>
  dplyr::summarise_at(bench_results, mean)

readr::write_csv(summary_by_datasets, here("data", "mean_bench_by_dataset.csv"))
```
## Generate plots using the mean of benchmark results for each datasets

The plots generated below is used for Figure 1 in the manuscript.

```{r}
# subset_apps <- c("AMAS (--check-align)", "goalign", "SEGUL (--datatype ignore)", "SEGUL", "AMAS")
subset_apps <- c("goalign",  "SEGUL", "AMAS", "SEGUL (--datatype ignore)")
subset_dataset <- c("Oliveros et al. 2019 (DNA)", "Jarvis et al. 2014 (DNA)", "Shen et al. 2018 (AA)")
summary_by_datasets |>
  dplyr::filter(Analyses != "Alignment splitting") |>
   dplyr::filter(OS_name == "Linux") |>
  dplyr::filter(Apps %in% subset_apps) |>
  dplyr::filter(Datasets %in% subset_dataset)|>
  ggplot(aes(x = Execution_time_secs, y = RAM_usage_Mb, color = Apps, shape = Datasets)) +
    geom_point( size = 3) +
    facet_wrap(~ Analyses, scales = "free", ncol = 2) +
    scale_colour_brewer(palette = "Dark2") +
    theme_classic() +
    labs(x = "Execution time (secs)", y = "Memory usage (Mb)")

ggsave(here("figures", "summary_fig1.png"), width = 9, height = 4, units = "in")
ggsave(here("figures", "summary_fig1.pdf"), width = 9, height = 4, units = "in")
```
## Re-plot for main text

```{r}
summary_by_datasets |>
  dplyr::filter(Analyses != "Alignment splitting") |>
  dplyr::filter(Apps != "Phyluce") |>
  ggplot(aes(x = Execution_time_secs, y = RAM_usage_Mb, color = Apps, shape = Datasets)) +
    geom_point( size = 2) +
    facet_wrap(~ Analyses + OS_name, scales = "free", ncol = 2) +
    scale_colour_brewer(palette = "Dark2") +
    theme_classic() +
    labs(x = "Execution time (secs)", y = "Memory usage (Mb)")

ggsave(here("figures", "summary_all.png"), width = 8, height = 8, units = "in")
ggsave(here("figures", "summary_all.pdf"), width = 8, height = 8, units = "in")
```


## Generate plots summary for Linux only

```{r}
summary_linux_dna <- latest_bench |>
  dplyr::filter(Analyses == "Alignment concatenation") |>
  dplyr::filter(OS_name == "Linux") |>
  dplyr::filter(Datatype == "DNA") |>
  dplyr::group_by(Analyses, Apps) |>
  dplyr::summarise_at(bench_results, mean)

summary_linux_dna |>
  ggplot(aes(x = Execution_time_secs, y = RAM_usage_Mb, color = Apps, shape = Apps)) +
    geom_point( size = 3) +
    scale_shape_manual(values=c(0,1,2,5,6,15,17)) +
    scale_colour_brewer(palette = "Dark2") +
    theme_classic() +
    labs(x = "Execution time (secs)", y = "RAM usage (Mb)")

ggsave(here("figures", "summary_linux.png"), width = 8, height = 4, units = "in")
ggsave(here("figures", "summary_linux.pdf"), width = 8, height = 4, units = "in")
```
## Randoms datasize check

```{r}
latest_bench |>
  dplyr::filter(Analyses == "Alignment concatenation") |>
  dplyr::filter(OS_name == "Linux") |>
  dplyr::filter(Pubs == "Oliveros et al. 2019") |>
  dplyr::group_by(Apps) |>
  dplyr::summarise_at(bench_results, mean)
  
```


## Generate bivariate plots for each benchmark result

```{r}
plot_bench <- function(.analyses, .os) {
  latest_bench |>
  dplyr::filter(Analyses == .analyses) |>
  dplyr::filter(OS_name == .os) |>
  dplyr::filter(Latest_bench == TRUE) |>
  ggplot(aes(x = Execution_time_secs, y = RAM_usage_Mb, color = Apps, shape = Datasets)) +
    geom_point() +
    theme_classic() +
    scale_colour_brewer(palette = "Set1") +
    labs(x = "Execution time (secs)", y = "Memory usage (Mb)")
}
```

## Names

```{r}
os_1 <- "macOS"
os_2 <- "Linux"
os_3 <- "Windows (WSL)"

fname <- function(os, analysis_name) {
  paste0(os, "_", analysis_name)
}
```


## Plot alignment summary

```{r}
stat_fname <- "summary_stats.png"

## Benchmark 1
plot_bench(.analyses = "Summary statistics", .os = os_1)
stat_fname_1 <- fname(os_1, stat_fname)
ggsave(here("figures", stat_fname_1), width = 8, height = 4, units = "in")

## Benchmark 2
plot_bench(.analyses = "Summary statistics", .os = os_2)
stat_fname_2 <- fname(os_2, stat_fname)
ggsave(here("figures", stat_fname_2), width = 8, height = 4, units = "in")

## Benchmark 3
plot_bench(.analyses = "Summary statistics", .os = os_3)
stat_fname_3 <- fname("WindowsWSL", stat_fname)
ggsave(here("figures", stat_fname_3), width = 8, height = 4, units = "in")
```

## Plot alignment concatenation

```{r}
concat_fname <- "concat_stats.png"

## Benchmark 1
plot_bench(.analyses = "Alignment concatenation", .os = os_1)
concat_fname_1 <- fname(os_1, concat_fname)
ggsave(here("figures", concat_fname_1), width = 8, height = 4, units = "in")

## Benchmark 2
plot_bench(.analyses = "Alignment concatenation", .os = os_2)
concat_fname_2 <- fname(os_2, concat_fname)
ggsave(here("figures", concat_fname_2), width = 8, height = 4, units = "in")

## Benchmark 3
plot_bench(.analyses = "Alignment concatenation", .os = os_3)
concat_fname_3 <- fname("WindowsWSL", concat_fname)
ggsave(here("figures", concat_fname_3), width = 8, height = 4, units = "in")
```

## Plot alignment splitting

```{r}
split_fname <- "split_alignment.png"

## Benchmark 1

## Benchmark 1
plot_bench(.analyses = "Alignment splitting", .os = os_2)
split_fname_2 <- fname(os_2, split_fname)
ggsave(here("figures", split_fname_2), width = 8, height = 4, units = "in")
```

